|

### 2021-07-22 prova_knit_finale.Rmd
aggiunto algoritmo
correzioni

Assume now that there exists a dominating measures $\nu(dy)$ such that 
$F_t(x_t,dy_t)=f_t(y_t\mid x_t)\nu(dy_t)$ for all $t.$
Then by the last equality of definition \ref{state-space-def}, the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{equation}\label{state-space-densities}
   \P_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T})=\P_T(dx_{0:T})\prod\limits_{t=0}^T f_t(y_t\mid x_t)\prod\limits_{t=0}^T\nu(dy_t).
\end{equation}
Integrating away $X_{0:T},$ one gets the marginal density
\begin{align*}\label{state-space-marginal-y}
  \P_T(dy_{0:T}) &= \E_{\P_{t}}\big[\prod\limits_{s=0}^t f_s(y_s\mid X_s) \big] \prod\limits_{s=0}^t\nu(dy_s) \\
                &= p_t(y_{0:t})\prod\limits_{s=0}^t\nu(dy_s),
\end{align*}                
where $p_t(y_{0:t})$ is the likelihood function.
Dividing the left-hand side of equation \ref{state-space-densities} by the left-hand side of equation \ref{state-space-marginal-y} and the right-hand side of equation \ref{state-space-densities} by the right-hand side of equation \ref{state-space-marginal-y}, one obtains the filtering distribution
\begin{equation}
\P_t(X_{0:t}\in dx_{0:t}\mid Y_{0:t}=y_{0:t})=\frac{1}{p_t(y_{0:t})}\big[\prod\limits_{s=0}^t f_s(y_s\mid X_s) \big]\P_t(dx_{0:t}).
\end{equation}

@@

Suppose that $\X=\R^{d_x}$ and $\Y=\R^{d_y}.$ Let $U_t\thicksim\N(0,\Sigma_t)$ and $V_t\thicksim\N(0,R_t)$ be two independent white-noise errors and assume the initial distribution 
$X_0\thicksim\N(0,\Sigma_0).$ We have a linear Gaussian model if 
\begin{align}
X_t &= A_tX_{t-1}+U_t \\
Y_t &= B_tX_t + V_t.
\end{align}
Note that  $X_t\mid X_{t-1}$ and $Y_t\mid X_t$ are both normally distributed. 
\
\begin{lemma}\label{kalman-filter}
  Let $X\thicksim\N(m_0,C_0)$ and $Y\mid X\thicksim\N(BX,R)$, where $R$ and $Q_0$ are positive semi-definite matrices. 
  Then, it holds that
  \begin{align}
    X\mid Y=y&\thicksim\N(m_1,C_1),\\
    C_1&= C_0 (I_{d_x} - B'(B C_0 B'+ R)^{-1} B C_0) \\
    m_1&=[I_{d_x} - C_0 B'(B C_0 B'+ R)^{-1} B ]m_0 + C_0 B'(B C_0 B'+ R)^{-1}y.
  \end{align}
\end{lemma}
\begin{proof*}
   The lemma can be proved using Bayes' Theorem. Suppose $X\thicksim\N(m_0, C_0).$ It is easy to see that
\begin{equation} \label{normal-distrib-rewrite}
        p(x)\propto \text{exp}\Big(-\frac{1}{2}X'C_0^{-1}X + X'C_0^{-1}m_0\Big).
\end{equation}
Now, by Bayes' Theorem,

    \begin{align*}
        p(x\mid Y=y)  &\propto p(x) p(y\mid x) \\
                    &\propto \text{exp}\Big(-\frac{1}{2}X' C_0^{-1}X + X' C_0^{-1}m_0-\frac{1}{2}(y-BX)'R^{-1}(y-BX)) \\
                    &\propto \text{exp}(-\frac{1}{2}X' C_0^{-1}X + X' C_0^{-1}m_0-\frac{1}{2}X' B' R^{-1}BX+X' B' R^{-1}y)\\
                    &=\text{exp}(-\frac{1}{2}X'(C_0^{-1}+B' R^{-1}B)X+
                    X'(C_0^{-1}m_0+B' R^{-1}y)\Big).
    \end{align*}

Hence, it follows that 
\begin{equation*}
    C_1=Var(X\mid Y=y)= (C_0^{-1}+ B' R^{-1} B)^{-1}=C_0-C_0 B'(B C_0 B'+R)^{-1}B C_0
\end{equation*}
and 
    \begin{align*}
        m_1 &=\E(X\mid Y=y) = C_1^{-1}(C_0^{-1}m_0+B' R^{-1}y)\\
            &= (C_0-C_0B'(B C_0B'+R)^{-1}B C_0)(C_0^{-1}m_0+B' R^{-1}y)\\
            &= m_0-C_0B'(B C_0B'+R)^{-1}B m_0 + C_0 B' R^{-1}y - C_0B'(B C_0B'+R)^{-1}B C_0 B' R^{-1}y \\
            &=[I_{d_x} - C_0 B'(B C_0 B'+ R)^{-1} B ]m_0 + C_0 B'(B C_0 B'+ R)^{-1}y.
    \end{align*}
\end{proof*}

Assume that $\P_{t-1}(dx_{t-1}\mid Y_{1:t-1}=y_{1:t-1})=\N(m_{t-1},C_{t-1}).$
Then, using the transition equation of the latent variable, 
by the linearity of the Gaussian distribution, we have that the predictive distribution is 
\begin{equation}
\P_{t-1}(dx_{t}\mid Y_{1:t-1}=y_{1:t-1})=\N(A_tm_{t-1},A_tC_{t-1}A_t'+\Sigma_t).
\end{equation}
Letting $E_T=A_tC_{t-1}A_t'+\Sigma_t,$ and applying lemma \ref{kalman-filter}, it follows that the filtering distribution is 
\begin{align*}
  \P_{t}(X_t\in dx_{t}\mid Y_{1:t-1}&=y_{1:t-1})=\N(m_t,C_t), \text{where} \\ C_t&=E_t[I_{d_x} - B_t'(B_tE_tB_t'+R_t)^{-1} B_tE_t] \\
  m_t &= [I_{d_x} - E_tB_t'(B_tE_tB_t'+R_t)^{-1} B_t]A_tm_{t-1}+E_tB_t'(B_tE_tB_t'+R_t)^{-1}y_t.
\end{align*}

>>>>

Assume now that there exists a dominating measures $\nu(dy)$ such that 
$F_t(x_t,dy_t)=f_t(y_t\mid x_t)\nu(dy_t)$ for all $t.$
Then by the last equality of definition \ref{state-space-def}, the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{equation}\label{state-space-densities}
   \P_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T})=\P_T(dx_{0:T})\prod\limits_{t=0}^T f_t(y_t\mid x_t)\prod\limits_{t=0}^T\nu(dy_t).
\end{equation}
Integrating away $X_{0:T},$ one gets the marginal density
\begin{equation}\label{state-space-marginal-y}
  \P_T(dy_{0:T}) = \E_{\P_{t}}\big[\prod\limits_{s=0}^t f_s(y_s\mid X_s) \big] \prod\limits_{s=0}^t\nu(dy_s)
                = p_t(y_{0:t})\prod\limits_{s=0}^t\nu(dy_s),
\end{equation}                
where $p_t(y_{0:t})$ is the likelihood function.
Dividing the left-hand side of equation \ref{state-space-densities} by the left-hand side of equation \ref{state-space-marginal-y} and the right-hand side of equation \ref{state-space-densities} by the right-hand side of equation \ref{state-space-marginal-y}, one obtains the distribution
\begin{equation}
\P_t(X_{0:t}\in dx_{0:t}\mid Y_{0:t}=y_{0:t})=\frac{1}{p_t(y_{0:t})}\big[\prod\limits_{s=0}^t f_s(y_s\mid X_s) \big]\P_t(dx_{0:t}).
\end{equation}

@@

Suppose that $\X=\R^{d_x}$ and $\Y=\R^{d_y}.$ Let $U_t\thicksim\N(0,\Sigma_t)$ and $V_t\thicksim\N(0,\Lambda_t)$ be two independent white-noise errors and assume the initial distribution 
$X_0\thicksim\N(0,\Sigma_0).$ We have a linear Gaussian model if 
\begin{align}
X_t &= A_tX_{t-1}+U_t \\
Y_t &= B_tX_t + V_t.
\end{align}
Note that  $X_t\mid X_{t-1}$ and $Y_t\mid X_t$ are both normally distributed. 
\
\begin{lemma}\label{kalman-filter}
  Let $X\thicksim\N(m_0,C_0)$ and $Y\mid X\thicksim\N(BX,\Lambda)$, where $R$ and $Q_0$ are positive semi-definite matrices. 
  Then, it holds that
  \begin{align}
    X\mid Y=y&\thicksim\N(m_1,C_1),\\
    C_1&= C_0 (I_{d_x} - B'(B C_0 B'+ \Lambda)^{-1} B C_0) \\
    m_1&=[I_{d_x} - C_0 B'(B C_0 B'+ \Lambda)^{-1} B ]m_0 + C_0 B'(B C_0 B'+ \Lambda)^{-1}y.
  \end{align}
\end{lemma}
\begin{proof*}
   The lemma can be proved using Bayes' Theorem. Suppose $X\thicksim\N(m_0, C_0).$ It is easy to see that
\begin{equation} \label{normal-distrib-rewrite}
        p(x)\propto \text{exp}\Big(-\frac{1}{2}x'C_0^{-1}x + x'C_0^{-1}m_0\Big).
\end{equation}
Now, by Bayes' Theorem,

    \begin{align*}
        p(x\mid Y=y)  &\propto p(x) p(y\mid x) \\
                    &\propto \text{exp}\Big(-\frac{1}{2}x' C_0^{-1}x + x' C_0^{-1}m_0-\frac{1}{2}(y-Bx)'\Lambda^{-1}(y-Bx)) \\
                    &\propto \text{exp}(-\frac{1}{2}x' C_0^{-1}x + x' C_0^{-1}m_0-\frac{1}{2}x' B' \Lambda^{-1}Bx+x' B' \Lambda^{-1}y)\\
                    &=\text{exp}(-\frac{1}{2}x'(C_0^{-1}+B' \Lambda^{-1}B)x+
                    x'(C_0^{-1}m_0+B' \Lambda^{-1}y)\Big).
    \end{align*}

Hence, it follows that 
\begin{equation*}
    C_1=Var(X\mid Y=y)= (C_0^{-1}+ B' \Lambda^{-1} B)^{-1}=C_0-C_0 B'(B C_0 B'+\Lambda)^{-1}B C_0
\end{equation*}
and 
\begin{equation*}
\begin{split}
        m_1 &=\E(X\mid Y=y) = C_1^{-1}(C_0^{-1}m_0+B' \Lambda^{-1}y)\\
            &= (C_0-C_0B'(B C_0B'+\Lambda)^{-1}B C_0)(C_0^{-1}m_0+B' \Lambda^{-1}y)\\
            &= m_0-C_0B'(B C_0B'+\Lambda)^{-1}B m_0 + C_0 B' \Lambda^{-1}y - C_0B'(B C_0B'+\Lambda)^{-1}B C_0 B' \Lambda^{-1}y \\
            &=[I_{d_x} - C_0 B'(B C_0 B'+ \Lambda)^{-1} B ]m_0 + C_0 B'(B C_0 B'+ \Lambda)^{-1}y.
\end{split}
\end{equation*}$\hfill\blacksquare$
\end{proof*}

Assume that $\P_{t-1}(dx_{t-1}\mid Y_{1:t-1}=y_{1:t-1})=\N(m_{t-1},C_{t-1}).$
Then, using the transition equation of the latent variable, 
by the linearity of the Gaussian distribution, we have that the predictive distribution is 
\begin{equation}
\P_{t-1}(dx_{t}\mid Y_{1:t-1}=y_{1:t-1})=\N(A_tm_{t-1},A_tC_{t-1}A_t'+\Sigma_t).
\end{equation}
Letting $R_T=A_tC_{t-1}A_t'+\Sigma_t,$ and applying lemma \ref{kalman-filter}, it follows that the filtering distribution is 
\begin{align*}
  \P_{t}(X_t\in dx_{t}\mid Y_{1:t-1}&=y_{1:t-1})=\N(m_t,C_t), \text{ where} \\ C_t&=R_t[I_{d_x} - B_t'(B_tR_tB_t'+\Lambda_t)^{-1} B_tR_t] \\
  m_t &= [I_{d_x} - R_tB_t'(B_tR_tB_t'+\Lambda_t)^{-1} B_t]A_tm_{t-1}+R_tB_t'(B_tR_tB_t'+\Lambda_t)^{-1}y_t.
\end{align*}
